---
title: "Refitting the Model on New Data"
format: pdf
editor: visual
---

```{r setup}
#| echo: false
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#"
)
```

```{r packages and setup}
#| include: false
#| echo: false

library("viridis")
library("tidymodels") 
theme_set(theme_bw() + 
            theme(panel.border = element_blank(), 
                  panel.grid.major = element_blank(),
                  panel.grid.minor = element_blank(), 
                  axis.line = element_line(colour = "black")))
options(ggplot2.continuous.colour="viridis")
library("extrafont")
library("tidypredict")
library("recipes")
library("knitr")
library("patchwork")
library("latex2exp")
```

### Data Preparation

The following code is the result of triage for what we thought was important for reproducibility of all the research that went into this paper. Other R scripts such as `analysis0.R`, `analysis1.R`, `analysis2.R` contain the messy, yet authentic code that was the outpouring of this modeling process. When someone is attempting to discover something they have quick ideas they don’t wanna forget so they just keep coding them up and their organization takes a backseat to discovery and inquiry.

### Data Preparation

```{r loading the data}
save <- FALSE

tuning_data <- read.csv(here::here("Data", "Interpolated_CompleteSet_May11_FlightData.csv"),
                         header = TRUE)

colnames(tuning_data) <- c("input_psd", "distance", "measured_psd")

tuning_data <- tuning_data |> mutate("transformed_distance" = distance^(-1/1000))
```

### Visualize Training Data
```{r visualize}
tuning_data |> 
  ggplot(aes(transformed_distance, measured_psd)) +
  geom_point(color = "#433E85FF", alpha = 0.03) +
  labs(title = TeX("Recieved PSD Over Distance$^{-0.001}$ (All Input PSD)"),
       caption = "Data plotted before any added noise with the Distance variable transformed.") +
  xlab(TeX("Distance$^{-0.001}$")) +
  ylab("Recieved PSD (dBW/200 MHz)") +
  theme(text=element_text(family="Times New Roman", size=14),
        plot.caption = element_text(hjust=0.5))
```

### Train/Test Split
```{r }
set.seed(613)
mod_split <- tuning_data |> 
  initial_split(
    prop = 0.8
  )

mod_test <- testing(mod_split)
mod_train <- training(mod_split)
```

### Model Fitting
```{r model fitting}
#| eval: true
m1 <- lm(measured_psd ~ input_psd + transformed_distance, 
         data = mod_train)

coef(m1) |> knitr::kable(col.names = c("Variable", "Coefficient Value"))
```

