---
title: "Predicting Aggregate Interference"
author: "Carson Slater"
format: pdf
editor: visual
---

```{r setup}
#| echo: false
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#"
)
```

```{r packages and setup}
#| include: false
#| echo: false

library("viridis")
library("tidymodels") 
theme_set(theme_bw() + 
            theme(panel.border = element_blank(), 
                  panel.grid.major = element_blank(),
                  panel.grid.minor = element_blank(), 
                  axis.line = element_line(colour = "black")))
options(ggplot2.continuous.colour="viridis")
library("extrafont")
# font_import() 
# loadfonts(device="pdf")
# fonts()
library("tidypredict")
library("recipes")
library("knitr")
library("patchwork")
library("latex2exp")
```

## Data Preparation and EDA

The following code is the result of triage for what we thought was important for reproducibility of all the research that went into this paper. Other R scripts such as `analysis0.R`, `analysis1.R`, `analysis2.R` contain the messy, yet authentic code that was the outpouring of this modeling process. When someone is attempting to discover something they have quick ideas they don’t wanna forget so they just keep coding them up and their organization takes a backseat to discovery and inquiry.

#### Data Preparation

```{r loading the data}
save <- FALSE

tuning_data <- read.csv(here::here("Data", "MamdaniTuningData_1-6.csv"),
                         header = TRUE)

distance <- as.vector(as.matrix(tuning_data[ ,c("tx1_distance",
                                                  "tx2_distance",
                                                  "tx3_distance")]))

input_psd <- as.vector(as.matrix(tuning_data[ ,c("tx1_psd",
                                                  "tx2_psd",
                                                  "tx3_psd")]))

scaled_psd <- as.vector(as.matrix(tuning_data[ ,c("tx1_scaled_psd",
                                                   "tx2_scaled_psd",
                                                   "tx3_scaled_psd")]))

aggregate_data <- cbind.data.frame(rep(tuning_data$rad_loc_idx, 3),
                                    distance,
                                    input_psd,
                                    scaled_psd)
```

We found it difficult to find real-world data for this kind of problem, so we have simulated data. To make our model robust, we decided to simulate random noise in our outcome of interest, `scaled_psd` (recieved power spectral density). We opted to assume a gaussian probability distribution with a mean of zero and a priori illicited variance. So to find the variance, we know the noise density (variance) at room temp is -174 dBm/Hz. We can convert to dBW/200MHz by using the following formula $$\sqrt{|-174 + 10\log_{10}(2\times10^8) - 30|} = 10.99953$$. We take the square root of the absolute value and then we have the standard deviation to parameterize the distribution of noise for the outcome of interest. So the noise is $\mathcal{N}(0, 10.99)$.

```{r data prep}
# generating potential transformation for the `distance` variable
# see analyis2.R for other transformations we tried
aggregate_data <- aggregate_data |> 
  mutate(transform5 = distance^(-1/1000))

# adding noise
set.seed(613)

sigma <- sqrt(abs(-174 + 10*log10(200e6) - 30))

noise <- rnorm(nrow(aggregate_data), 0, sigma)

noise_aggregate_data <- aggregate_data

noise_aggregate_data$scaled_psd <- noise_aggregate_data$scaled_psd + noise
```

#### Data Visualization

In this our exploratory data analysis, I observed there is a non-linear relationship between `distance` and the `scaled_psd`.

```{r plots}
viridis_pal(option = "D")(12)

# No noise
if (save == TRUE) pdf(file = "Figures/rawdata.pdf", height = 5, width = 7)
aggregate_data |> 
  ggplot(aes(distance, scaled_psd)) +
  geom_point(color = "#433E85FF", alpha = 0.03) +
  labs(title = "Recieved PSD Over Distance Without Noise (All Input PSD)",
       caption = "Raw simulated data plotted before any added noise without the distance variable transformed.") +
  xlab(TeX("Distance (m)")) +
  ylab("Recieved PSD (dBW/200 MHz)") +
  theme(text=element_text(family="Times New Roman", size=14),
        plot.caption = element_text(hjust=0.5))
if (save == TRUE) dev.off()

 # Jitter (noise) added
if (save == TRUE) pdf(file = "Figures/noise_rawdata.pdf", height = 5, width = 7)
noise_aggregate_data |> 
  ggplot(aes(distance, scaled_psd)) +
  geom_point(color = "#433E85FF", alpha = 0.03) +
  labs(title = "Recieved PSD Over Distance With Noise (All Input PSD)",
       caption = "Raw simulated data plotted with simulated noise without the distance variable transformed.") +
  xlab("Distance (m)") +
  ylab("Recieved PSD (dBW/200 MHz)") +
  theme(text=element_text(family="Times New Roman", size=14),
        plot.caption = element_text(hjust=0.5))
if (save == TRUE) dev.off()

# No noise + transformed distance
if (save == TRUE) pdf(file = "Figures/tfdata.pdf", height = 5, width = 7)
aggregate_data |> 
  ggplot(aes(transform5, scaled_psd)) +
  geom_point(color = "#433E85FF", alpha = 0.03) +
  labs(title = TeX("Recieved PSD Over Distance$^{-0.001}$ (All Input PSD)"),
       caption = "Data plotted before any added noise with the Distance variable transformed.") +
  xlab(TeX("Distance$^{-0.001}$")) +
  ylab("Recieved PSD (dBW/200 MHz)") +
  theme(text=element_text(family="Times New Roman", size=14),
        plot.caption = element_text(hjust=0.5))
if (save == TRUE) dev.off()

# Noise + transformed distance
if (save == TRUE) pdf(file = "Figures/noise_tfdata.pdf", height = 5, width = 7)
noise_aggregate_data |> 
  ggplot(aes(transform5, scaled_psd)) +
  geom_point(color = "#433E85FF", alpha = 0.03) +
  labs(title = TeX("Recieved PSD Over Distance$^{-0.001}$ (All Input PSD)"),
       caption = "Data plotted after adding noise and transforming the distiance variable.") +
  xlab(TeX("Distance$^{-0.001}$")) +
  ylab("Recieved PSD (dBW/200 MHz)") +
  theme(text=element_text(family="Times New Roman", size=14),
        plot.caption = element_text(hjust=0.5))
if (save == TRUE) dev.off()
```

## Modeling

We create a `parsnip` linear model object.

```{r}
lm_mod <-
  parsnip::linear_reg() |>
  parsnip::set_engine("lm") |>
  parsnip::set_mode("regression")
```

#### The Model (Without Noise)

```{r m1 training}
set.seed(613)
mod_split <- aggregate_data |> 
  initial_split(
    prop = 0.8
  )

mod_test <- testing(mod_split)
mod_train <- training(mod_split)

# `tidymodels` Procedure for Fitting Linear Model.
my_recipe <- recipe(scaled_psd ~ input_psd + transform5, 
                    data = aggregate_data)

baked_recipe <- bake(prep(my_recipe), new_data = NULL)

lm_wkflow <- workflow() |> 
  add_model(lm_mod) |> 
  add_recipe(my_recipe)

# Specifying Model Metrics
my_metrics <- metric_set(mape, rsq, rmse)

# Fit the model (not with v-fold CV)
lm_fit <- fit(lm_wkflow, 
              data = mod_train)

lm_pred_test <- augment(lm_fit, 
                        new_data = mod_test)

my_metrics(lm_pred_test, 
           truth = scaled_psd, 
           estimate = .pred)

# THE MODEL ---------------------------------------------------------------
lm_fit |> 
  tidy()

m1 <- lm(scaled_psd ~ input_psd + transform5, 
         data = mod_train)
summary(m1)
```

#### Cross Validation (No Noise)

```{r}
# Creating CV Folds to Confirm Model Metrics
lm_folds <- vfold_cv(mod_train, v = 10, repeats = 5)

# Fitting the Model (with CV)

lm_folds_fitted <- lm_wkflow |>
  fit_resamples(resamples = lm_folds, 
                metrics = my_metrics)

# collect_metrics(lm_folds_fitted, summarize = F) |> 
#   filter(.metric == "mape") |> 
#   arrange(desc(.estimate)) |>
#   kable()

collect_metrics(lm_folds_fitted, summarize = T) |>
  select(.metric, mean, n, std_err) |>
  kable()
```

#### Benchmarking

```{r}
# Benchmark ---------------------------------------------------------------
columns <- c("Expression",
             "Min. Runtime",
             "Med. Runtime",
             "Itr/Sec",
             "Memory Allocated")

bnch <- bench::mark(tidymodels = lm_wkflow |> 
                      fit_resamples(resamples = lm_folds, metrics = my_metrics),
                    check = FALSE)
bnch |> 
  select(expression, min, 
         median, `itr/sec`, 
         mem_alloc) |> 
  knitr::kable(digits = 3, 
               col.names = columns) |> 
  kableExtra::kable_styling(font_size = 10,
                            latex_options = "HOLD_position") 
```

#### The Model (With Noise)

```{r m2 training}
# Performing Model Training

set.seed(613)
noise_mod_split <- noise_aggregate_data |> 
  initial_split(
    prop = 0.8
  )

noise_mod_test <- testing(noise_mod_split)
noise_mod_train <- training(noise_mod_split)

# `tidymodels` Procedure for Fitting Linear Model.
noise_my_recipe <- recipe(scaled_psd ~ input_psd + transform5, 
                    data = aggregate_data)

noise_baked_recipe <- bake(prep(my_recipe), new_data = NULL)

noise_lm_wkflow <- workflow() |> 
  add_model(lm_mod) |> 
  add_recipe(noise_my_recipe)

# Specifying Model Metrics
noise_my_metrics <- metric_set(mape, rsq, rmse)

# Fit the model (not with v-fold CV)
noise_lm_fit <- fit(lm_wkflow, 
              data = noise_mod_train)

noise_lm_pred_test <- augment(noise_lm_fit, 
                        new_data = noise_mod_test)

noise_my_metrics(noise_lm_pred_test, 
           truth = scaled_psd, 
           estimate = .pred)

# THE MODEL
noise_lm_fit |> 
  tidy()

m2 <- lm(scaled_psd ~ input_psd + transform5, 
         data = noise_mod_train)
summary(m2)
```

#### Cross Validation (With Noise)

```{r}
# Creating CV Folds to Confirm Model Metrics
noise_lm_folds <- vfold_cv(noise_mod_train, v = 10, repeats = 5)

# Fitting the Model (with CV)

noise_lm_folds_fitted <- noise_lm_wkflow |>
  fit_resamples(resamples = noise_lm_folds, 
                metrics = my_metrics)

# collect_metrics(noise_lm_folds_fitted, summarize = F) |> 
#   filter(.metric == "mape") |> 
#   arrange(desc(.estimate)) |>
#   kable()

collect_metrics(noise_lm_folds_fitted, summarize = T) |>
  select(.metric, mean, n, std_err) |>
  kable()
```

#### Benchmarking

```{r}
# Benchmark ---------------------------------------------------------------
columns <- c("Expression",
             "Min. Runtime",
             "Med. Runtime",
             "Itr/Sec",
             "Memory Allocated")

bnch <- bench::mark(tidymodels = noise_lm_wkflow |> 
                      fit_resamples(resamples = noise_lm_folds, metrics = my_metrics),
                    check = FALSE)
bnch |> 
  select(expression, min, 
         median, `itr/sec`, 
         mem_alloc) |> 
  knitr::kable(digits = 3, 
               col.names = columns) |> 
  kableExtra::kable_styling(font_size = 10,
                            latex_options = "HOLD_position") 
```

## Model Plotting

#### Making Confidence/Prediction Intervals

```{r}
n <- nrow(lm_pred_test)
Sxx <- sum((aggregate_data$transform5 - mean(aggregate_data$transform5))^2)
xstar <- seq(min(aggregate_data$transform5), max(aggregate_data$transform5), len = 71070)

sigma_m1 <- summary(m1)$sigma
sigma_m2 <- summary(m2)$sigma

ci_m1 <- qt(1 - 0.05/2, n - 2)*sigma_m1*sqrt(1/n + (xstar - mean(aggregate_data$transform5))^2/Sxx)
ci_m2 <- qt(1 - 0.05/2, n - 2)*sigma_m2*sqrt(1/n + (xstar - mean(noise_aggregate_data$transform5))^2/Sxx)

pi_m1 <- qt(1 - 0.05/2, n - 2)*sigma_m1*sqrt(1 + 1/n + (xstar - mean(aggregate_data$transform5))^2/Sxx)
pi_m2 <- qt(1 - 0.05/2, n - 2)*sigma_m2*sqrt(1 + 1/n + (xstar - mean(noise_aggregate_data$transform5))^2/Sxx)
```

```{r}
# no noise
# predicting on the whole data set
pred <- augment(lm_fit, new_data = aggregate_data)
pred <- pred |> mutate(upper_CI = .pred + ci_m1,
                       lower_CI = .pred - ci_m1,
                       upper_PI = .pred + pi_m1,
                       lower_PI = .pred - pi_m1)

# with noise
# predicting on the whole data set
noise_pred <- augment(noise_lm_fit, new_data = noise_aggregate_data)
noise_pred <- noise_pred |> mutate(upper_CI = .pred + ci_m2,
                                   lower_CI = .pred - ci_m2,
                                   upper_PI = .pred + pi_m2,
                                   lower_PI = .pred - pi_m2)
ipsd55 <- noise_pred |> 
  filter(input_psd == -55)

# visualizing the models
if (save == TRUE) pdf(file = "Figures/pred.pdf", height = 5, width = 7)
pred |> 
  filter(input_psd == -55) |> 
  ggplot(aes(transform5, scaled_psd)) +
    geom_point(color = "#433E85FF", alpha = 0.5, size = 3) +
    geom_line(aes(transform5, .pred), color = "darkorange", lwd = 1) +
    geom_line(aes(transform5, upper_PI), color = "darkorange", linetype = "dotted", lwd = 1) +
    geom_line(aes(transform5, lower_PI), color = "darkorange", linetype = "dotted", lwd = 1) +
    labs(title = "Regression Model Fit (Transmitted PSD = -55)",
         caption = "Model fit and prediction interval (no noise added).") +
    xlab(TeX("Distance$^{-0.001}$")) +
    ylab("Recieved PSD (dBW/200 MHz)") +
    theme(text=element_text(family="Times New Roman", size=14),
          plot.caption = element_text(hjust=0.5))
if (save == TRUE) dev.off()

if (save == TRUE) pdf(file = "Figures/noise_pred.pdf", height = 5, width = 7)
noise_pred |> 
  filter(input_psd == -55) |> 
  ggplot(aes(transform5, scaled_psd)) +
    geom_point(color = "#433E85FF", alpha = 0.2, size = 3) +
    geom_line(aes(transform5, .pred), color = "darkorange", lwd = 1) +
    geom_line(aes(transform5, upper_PI), color = "darkorange", linetype = "dotted", lwd = 1) +
    geom_line(aes(transform5, lower_PI), color = "darkorange", linetype = "dotted", lwd = 1) +
    labs(title = "Regression Model Fit (Transmitted PSD = -55)",
         caption = "Model fit and prediction interval with training data (noise added).") +
    xlab(TeX("Distance$^{-0.001}$")) +
    ylab("Recieved PSD (dBW/200 MHz)") +
    theme(text=element_text(family="Times New Roman", size=14),
          plot.caption = element_text(hjust=0.5))
if (save == TRUE) dev.off()
```


### Model Diagnostics
```{r}
#| eval: FALSE
ri <- rstandard(m1)

qqnorm(ri); qqline(ri)

plot(predict(m1, newdata = mod_train), ri, pch = 19, col = alpha(1, 0.05))
abline(h = 0, col = "red")
```
